Analysis for Biracial Grade Retention Project

### 2022-09-02 analysis/analysis.Rmd
Add code chunks to create dummied dataset, run models, and test margins

>>>>

```{r create-dummied-data}
acs_dummied <- acs %>%
  # first we need to unorder current grade for dummy_cols to work
  mutate(current_grade=factor(current_grade, levels=levels(current_grade), 
                              ordered = FALSE)) %>%

  dummy_cols(remove_first_dummy = TRUE) %>%
  select(cluster, strata, perwt, below_exp_grade, year, family_income, 
         parents_married, own_home, degree_mother, degree_father,
         foreign_born, foreign_born_mother, foreign_born_father, 
         spk_eng, spk_eng_father, spk_eng_mother,
         starts_with(c("state_","current_grade_","race","sex_","metro_"))) %>%
  mutate(year=year-2010) %>%
  clean_names()
```

```{r model-dummied}
model_full <- glm(below_exp_grade~current_grade_1st*year+current_grade_2nd*year+
                    current_grade_3rd*year+current_grade_4th*year+
                    current_grade_5th+current_grade_6th*year+
                    current_grade_7th*year+current_grade_8th*year+
                    current_grade_9th*year+current_grade_10th*year+
                    current_grade_11th*year+current_grade_12th*year+
                    race_black+race_indigenous+race_asian+race_latino+
                    race_black_white+race_black_indigenous+race_black_latino+
                    race_black_asian+race_white_indigenous+race_white_latino+
                    race_white_asian+race_indigenous_latino+
                    race_indigenous_asian+race_latino_asian+
                    sex_female+
                    metro_non_metro+metro_central_city+metro_metro_non_central+
                    metro_metro_indeterminable+state_arizona+state_arkansas+
                    state_california+state_colorado+state_connecticut+state_dc+
                    state_delaware+state_florida+state_georgia+state_hawaii+
                    state_idaho+state_illinois+state_indiana+state_iowa+
                    state_kansas+state_kentucky+state_louisiana+state_maine+
                    state_maryland+state_massachusetts+state_michigan+
                    state_minnesota+state_mississippi+state_missouri+
                    state_montana+state_nebraska+state_nevada+
                    state_new_hampshire+state_new_jersey+state_new_mexico+
                    state_new_york+state_north_carolina+state_north_dakota+
                    state_ohio+state_oklahoma+state_oregon+state_pennsylvania+
                    state_rhode_island+state_south_carolina+state_south_dakota+
                    state_tennessee+state_texas+state_utah+state_vermont+
                    state_virginia+state_washington+state_west_virginia+
                    state_wisconsin+state_wyoming+
                    +sqrt(family_income/10000)+parents_married+own_home+
                    degree_mother+degree_father+
                    spk_eng+spk_eng_mother+spk_eng_father+
                    foreign_born_mother+foreign_born_father+foreign_born, 
                  data=acs_dummied, family=binomial)
deviance(model_full)
# FIXME - its a little higher so I am missing something
```

```{r}
# what needs to get set at mean?
# - current_grade
# - sex
# - race
# - state
# - metro
# I will not do this for year because I want to use that to grab the modal
# year of 2014
means <- acs_dummied %>%
  select(year, starts_with(c("state_","current_grade_","race_","sex_","metro_")))
means <- lapply(means, function(x) {
  rep(mean(x), length(x))
}) %>%
  bind_rows()

cf_df <- acs_dummied %>%
  select(below_exp_grade, race,family_income, parents_married, own_home, 
         degree_mother, degree_father, foreign_born, foreign_born_mother, 
         foreign_born_father, spk_eng, spk_eng_father, spk_eng_mother) %>%
  bind_cols(means)
```

```{r}
# I think I can use year in marg to get the results for the modal year
temp <- subset(cf_df, race=="Black/White")
#temp$year <- 4
lor <- predict(model_full, newdata=temp)
mean(exp(lor)/(1+exp(lor)))
marg(model_full, "year", data=subset(cf_df, race=="Black/White"))
marg(model_full, "year", data=subset(cf_df, race=="White"))
marg(model_full, "year", data=subset(cf_df, race=="Black"))
```
